
\chapter{ATU Robotics Club Constitution}
(Reformatted for notebook purposes.)

\section{Charter}
\subsection{1.}
The official name of this organization will be \textit{The Arkansas Tech University Robotics Club}.
\subsection{2.} The purpose of this organization is to:
\begin{enumerate}
    \item Participate in competitive robotics programs.
    \item Practice quality scientific and engineering principles.
    \item Promote STEAM education throughout Arkansas.
\end{enumerate}

\section{Membership}
\subsection{1.} Members of the Robotics Club must be collegiate students;
neither hailing university nor major will prevent membership.
\subsection{2.} Members will not be discriminated against on undue bases
including but not limited to: race, gender, sexual orientation, nationality,
veteran status, and disability.
\subsection{3.} Members of the Robotics Club must be in good standing with
this constitution and club ordinances.
\subsubsection{1.} Members found in violation of this section
are subject to removal from the club at the discretion of the club
officers and advisor.

\section{Officers}
\subsection{1.} The office of \textit{President} is established with the following
duties:
\begin{enumerate}
    \item Acting as a representative for the club at events and in media.
    \item Planning and guiding meetings.
    \item Monitoring and guiding ATU Robotics programs.
    \item Acting as a mediator and final say in club decisions.
    \item Encouraging quality engineering principles.
    \item Delegating officers as beneficial for ATU Robotics.
\end{enumerate}
\subsection{2.} The office of \textit{Vice President} is established with the following duties:
\begin{enumerate}
    \item Aiding in the duties of the President.
    \item Acting as a replacement for other officers in their absence.
\end{enumerate}
\subsection{3.} The office of \textit{Secretary} is established with the following duties:
\begin{enumerate}
    \item Keeping record during club events.
    \item Ensuring club information is easily accessible and known to members.
    \item Working with the Treasurer to manage logistics for club events.
    \item Working with the Treasurer to manage club resources.
\end{enumerate}
\subsection{4.} The office of \textit{Treasurer} is established with the following duties:
\begin{enumerate}
    \item Managing the finances of ATU Robotics.
    \item Managing fundraising programs.
    \item Working with the Secretary to manage logistics for club events.
    \item Working with the Secretary to manage club resources.
\end{enumerate}
\subsection{5.} The office of \textit{Social Relations Manager} is established with the following duties:
\begin{enumerate}
    \item Managing public outreach.
    \item Managing social media.
\end{enumerate}
\subsection{6.} Officers will be elected by simple majority of present club
members at the first possible meeting of the year or immediately following
an officer's removal.
\subsection{7.} Officer terms will be no more than one year unless re-elected.
\subsection{8.} Officers may hold only one position.
\subsection{9.} Any club member may propose to remove an officer at a meeting;
at which point a simple majority vote is required for removal.
\subsection{10.} All officers are expected to be active in the club and exemplify
the principles of ATU Robotics.
\subsection{11.} The President may establish specialized \textit{Executive Officer}
positions as beneficial to club operations.
\subsubsection{1.} This position will last no longer than the remaining term of
the President.
\subsubsection{2.} Executive Officer positions act as an exception to Article 3,
section{} 7.
\subsection{12.} The Treasurer may propose \textit{Membership Dues} at meetings which will be
levied with the approval of the President.
\subsubsection{1.} Extenuating circumstances may excuse a member from paying dues while
remaining in the club.

\section{Operations}
\subsection{1.} Meeting dates are tentative and decided with regard to the schedules of members at
the beginning of each year.
\subsection{2.} This constitution and its version history will be easily available to members
and examined upon each election.
\subsection{3.} Major changes may be proposed by any member at a meeting and decided by two-thirds
majority vote of present members.
\subsubsection{1.} Major changes to this document include adding, removing, or significantly editing sections.
\subsection{4.} Minor changes may be applied immediately with no vote necessary.
\subsubsection{1.} Minor changes to this document include stylistic changes, grammar corrections, and insignificant
clarification.